---
title: "AccRQA: Accelerated Recurrence Quantification Analysis"
output: rmarkdown::html_vignette
author: "Karel Adámek, Jan Novotný"
vignette: >
  %\VignetteIndexEntry{AccRQA: Accelerated Recurrence Quantification Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(AccRQA)
```

## Introduction

**AccRQA** is a multi-platform, high-performance library for computing
*Recurrence Quantification Analysis* (RQA) measures.
The library supports execution on standard CPUs as well as on
NVIDIA GPUs, with GPU acceleration being entirely optional.
An NVIDIA GPU is therefore **not required** to use AccRQA.

AccRQA provides optimized implementations of core RQA measures and
exposes them through consistent interfaces in **R, Python, and C/C++**.
The design goal of the library is to enable efficient large-scale RQA
workflows while preserving numerical consistency across platforms.

A key feature of AccRQA is its **platform-agnostic interface**.
The choice of computational backend (CPU or GPU) is controlled by a
single parameter in the function invocation.
If the requested backend is unavailable on the current system,
AccRQA automatically falls back to CPU execution.
This design allows users to seamlessly transition from local
workflows (e.g., laptops or workstations) to high-performance
computing (HPC) environments equipped with GPUs, without modifying
analysis code.

Currently supported computational platforms include:

- CPUs (with optional multi-threading)
- NVIDIA GPUs (Linux and macOS)


All AccRQA interfaces assume that input data initially reside in
main memory (CPU RAM).
When GPU execution is selected, the data are transferred internally
to the appropriate device.

---

## Scope of the library

The AccRQA library exposes a compact and focused API that covers the
most commonly used RQA measures.
Specifically, the following quantities can be computed:

- **Recurrence Rate (RR)**
- **Laminarity (LAM)**, including trapping time (TT) and maximum trapping time (TT_max)
- **Determinism (DET)**, including average diagonal line length (L),
  maximum diagonal line length (L_max), and entropy (ENTR)

Distance computations currently support the **Euclidean norm** and the
**maximum norm**, with additional metrics planned for future
versions.

---

### Notes on performance and validation

The functions performs internal validation of input lengths and parameter
values to ensure consistency of the reconstructed phase space.
All computations are delegated to a compiled backend, making the
function suitable for large parameter sweeps and high-throughput
workflows.

GPU execution can significantly reduce computation time for large
inputs, but is not required for correct usage of the function.

---

# Contributing to AccRQA

People can contribute in different ways — code, documentation, bug reports, suggesting features or by creating examples and tutorials.

---

## Reporting Bugs

If you find a bug, please:

1. Check if it’s already reported.
2. If not, [open a new issue](https://github.com/KAdamek/AccRQA/issues).
3. Include:
   - Clear title and description
   - Steps to reproduce
   - Expected vs actual behavior
   - Environment info (Python version, OS, etc.)

---

## Suggesting Enhancements

If you miss a feature or you have an idea how to improve AccRQA package you can tell us about it using github issues.

1. Check for similar feature requests.
2. [Open a new issue](https://github.com/KAdamek/AccRQA/issues)
3. Describe:
   - What the feature does
   - Why it’s useful
   - Any alternatives you’ve considered

---

## Repository contributions

If you are happy to contribute to the code, you can do so via git branches and pull requests. Contributions can be made in Python or R interfaces and functions as well as in C/C++ library. The C/C++ library contains CPU functions parallelised using OpenMP and CUDA GPU kernels. Further work could be done to optimise the calculation of RQA for specific use cases and improve existing performance.


## License
By contributing, you agree that your contributions will be licensed under the same license as the project. See LICENSE for more info.


## Conclusion

AccRQA provides a unified and efficient framework for recurrence quantification analysis across a range of computational platforms. Its platform-agnostic interface and compiled backends make it suitable for both exploratory analysis on local machines and large-scale studies on HPC systems.



